\chapter{Survival Analysis}

\section{General Formulation}

\begin{definition}[Survival Function]
	The survival function\footnote{The survival function is the probability that the time of death is later than some specified time \(t\).} is defined to be
	\begin{equation}
		S(t)=P(T>t)=\int_{t}^{\infty}f(u)\dif u=1-F(t).
	\end{equation}
	where \(t\) is some specified time, \(T\) is a random variable denoting the time of death.
\end{definition}

\begin{definition}[Lifetime Distribution Function]
	The lifetime distribution function is defined to be
	\begin{equation}
		F(t)=P(T\leq t)
	\end{equation}
	If \(F\) is differentiable then the derivative, which is the density function of the lifetime distribution\footnote{The function \(f\) is sometimes called the event density; it is the rate of death or failure events per unit time.}, is defined to be
	\begin{equation}
		f(t)=F^{\prime}(t)=\frac{\dif}{\dif t}F(t)
	\end{equation}
\end{definition}

\begin{definition}[Hazard Function]
	The Hazard function\footnote{The Hazard function is the event rate at time \(t\) conditional on survival until time \(t\) or later (that is, \(T\geq t\)).} is defined to be
	\begin{equation}
		\lambda(t)=\lim_{\varepsilon\rightarrow 0^{+}}\left[\frac{P(t\leq T<t+\varepsilon\mid T\geq t)}{\varepsilon}\right]=\frac{f(t)}{S(t)}
	\end{equation}
\end{definition}

\begin{property}
	The relationship among \(\lambda(t)\), \(f(t)\), \(S(t)\),
	\begin{enumerate}
		\item
		      \begin{equation}
			      \lambda(t)=-\frac{\dif\log [S(t)]}{\dif t}
		      \end{equation}
		\item
		      \begin{equation}
			      S(t)=\exp\left[-\int_{0}^{t}\lambda(x)\dif x\right]
		      \end{equation}
		\item
		      \begin{equation}
			      f(t)=\lambda(t)\exp\left[-\int_{0}^{t}\lambda(x)\dif x\right]
		      \end{equation}
	\end{enumerate}
\end{property}

\begin{proof}

\end{proof}

\begin{example}[Constant Hazards]
	Suppose
	\begin{equation}
		\lambda(t)=\lambda
	\end{equation}
	then
	\begin{gather*}
		S(t)=\exp\left[-\int_{0}^{t}\lambda(x)\dif x\right]=\exp\left[-\int_{0}^{t}\lambda\dif x\right]=\exp(-\lambda t) \\
		f(t)=\lambda(t)\exp\left[-\int_{0}^{t}\lambda(x)\dif x\right]=\lambda\exp\left[-\int_{0}^{t}\lambda\dif x\right]=\lambda\exp(-\lambda t)
	\end{gather*}
	which is the exponential distribution.
\end{example}

\begin{example}[Bathtub Hazards]
	\begin{equation}
		\lambda(t)=\alpha t+\frac{\beta}{1+\gamma t}
	\end{equation}
\end{example}

\section{Estimation of Survival Function}

\paragraph*{Parametric Approach}

Suppose \(t_{1},t_{2},\ldots,t_{n}\) are failure times corresponding to censor indicators \(\delta_{1},\delta_{2},\ldots,\delta_{n}\). The likelihood function is
\begin{equation}
	\begin{aligned}
		f\left(\boldsymbol{\theta}\mid\mathbf{t},\boldsymbol{\delta}\right)= & \prod_{i=1}^{n}\left[f\left(t_{i}\right)\right]^{\delta_{i}}\left[S\left(t_{i}\right)\right]^{1-\delta_{i}} \\
		=                                                                    & \prod_{i=1}^{n}\left(\frac{f\left(t_{i}\right)}{S\left(t_{i}\right)}\right)^{\delta_{i}}S\left(t_{i}\right) \\
		=                                                                    & \prod_{i=1}^{n}\left[\lambda\left(t_{i}\right)\right]^{\delta_{i}}S\left(t_{i}\right)
	\end{aligned}
\end{equation}
where \(\lambda(t),S(t)\) depends on some parameter \(\theta\).

\begin{example}
	Suppose \(\boldsymbol{T}\) have exponential density, that,
	\begin{equation*}
		f(t)=\lambda \mathrm{e}^{-\lambda t},\quad S(t)=\mathrm{e}^{-\lambda t}
	\end{equation*}
	Thus,
	\begin{equation*}
		\begin{aligned}
			\ell(\lambda)= & \log[\ell(\theta)]=\sum_{i=1}^{n}\left[\delta_{i}\log(\lambda)-\lambda t_{i}\right]        \\
			=              & \left(\sum_{i=1}^{n}\delta_{i}\right)\log(\lambda)-\lambda\left(\sum_{i=1}^{n}t_{i}\right)
		\end{aligned}
	\end{equation*}
	Hence,
	\begin{equation*}
		\frac{\partial\ell(\lambda)}{\partial\lambda}=\frac{\sum_{i=1}^{n}\delta_{i}}{\lambda}-\sum_{i=1}^{n}t_{i}=0\Rightarrow\hat{\lambda}=\frac{\sum_{i=1}^{n}\delta_{i}}{\sum_{i=1}^{n}t_{i}}
	\end{equation*}
\end{example}

\paragraph*{Nonparametric Approach}

Then, for \(t_{(k)}\leq t<t_{(k+1)}\),
\begin{equation}
	\begin{aligned}
		\hat{S}(t)= & \prod_{j=1}^{k}\left(\frac{n_{j}-d_{j}}{n_{j}}\right)                                                                                             \\
		=           & \left(1-\frac{d_{1}}{n_{1}}\right)\left(1-\frac{d_{2}}{n_{2}}\right) \cdots\left(1-\frac{d_{k}}{n_{k}}\right)                                     \\
		\approx     & \left[1-\hat{\lambda}\left(t_{1}\right)\right]\left[1-\hat{\lambda}\left(t_{2}\right)\right] \ldots\left[1-\hat{\lambda}\left(t_{k}\right)\right]
	\end{aligned}
\end{equation}
where \(\hat{S}(t)\) is referred to as Kaplan-Meier estimate.

\section{Proportional Hazards Model}

Let \(t_{1},t_{2},\ldots,t_{n}\) be the failure times associated with censor indicator \(\delta_{1},\delta_{2},\ldots,\delta_{n}\) and the covariate vectors \(\bfx_{i}\).

Further, let \(t_{(1)}\leq t_{(2)}\leq\cdots\leq t_{(m)}\) be the ordered uncensored failure times corresponding to \(\delta_{(j)}=1,j=1,2,\ldots,m\), and \(x_{(1)},x_{(2)},\ldots,x_{(m)}\) are the associated covariate vectors. Note \((j)\) represents the label for the individual who dies at \(t_{(j)}\).

The proportional hazards model specifying the hazard at time \(t\) for an individual whose covariate vector is \(\bfx\) is given by
\begin{equation}
	\lambda(t)=\lambda_{0}(t)\mathrm{e}^{\bfx^{\top}\bfbeta}
\end{equation}
where \(\lambda_{0}(t)\) is referred to as the baseline hazard function.

The exact likelihood function is
\begin{equation}
	\ell\left[\beta,\lambda_{0}(t)\right]=\prod_{i=1}^{n}\left[\lambda_{i}\left(t_{i}\right)\right]^{\delta_{i}}S\left(t_{i}\right)
\end{equation}
depends on both the nonparametric function \(\lambda_{0}(t)\) and the parameter \(\bfbeta\). Thus, it might be difficult to estimate \(\lambda_{0}(t)\) and \(\bfbeta\) simultaneously.

The partial likelihood function is
\begin{equation}
	\ell_{p}(\bfbeta)=\prod_{j=1}^{m}\frac{\mathrm{e}^{\bfx_{(j)}^{\prime}\bfbeta}}{\sum_{l\in R\left(t_{(j)}\right)}\mathrm{e}^{\bfx_{l}^{\prime}\bfbeta}}=\prod_{i=1}^{n}\left[\frac{\mathrm{e}^{\bfx_{i}^{\top}\bfbeta}}{\sum_{l\in R\left(t_{i}\right)}\mathrm{e}^{\bfx_{l}^{\prime}\bfbeta}}\right]^{\delta_{i}}
\end{equation}
where \(R(t)\) is the set of individuals who are alive and uncensored at a time just before \(t_{i}\), which is called the risk set.

Another set of work converts the survival prediction problem into a classification problem by dividing the continuous time-to-event into nonoverlapping intervals

\section{Accelerated Failure Time Model}

The Accelerated Failure Time (AFT) model posits that the logarithm of the survival time \( T \) is linearly related to a set of covariates \( \bfX \):

\begin{equation}
	\log T = -\bfX^\top \bfbeta + \varepsilon,
\end{equation}
where \( \varepsilon \) is a random error term whose distribution determines the baseline survival characteristics. By exponentiating both sides, we obtain

\begin{equation*}
	T = \exp(-\bfX^\top \bfbeta + \varepsilon) = \exp(\varepsilon) \cdot \exp(-\bfX^\top \bfbeta).
\end{equation*}

Let \( T_0 = \exp(\varepsilon) \) denote the baseline survival time, with survival function \( S_0(t) = \bbP(T_0 > t) \). The conditional survival function under the AFT model is then given by
\begin{equation*}
	\begin{aligned}
		S(t|\bfX) & = \bbP(T > t \mid \bfX)                                      \\
		          & = \bbP(\exp(\varepsilon) \cdot \exp(-\bfX^\top \bfbeta) > t) \\
		          & = \bbP\left(\varepsilon > \log t + \bfX^\top \bfbeta \right) \\
		          & = S_0(t \exp(\bfX^\top \bfbeta)).
	\end{aligned}
\end{equation*}

The corresponding cumulative hazard function is
\begin{equation*}
	\Lambda_{\bfX}(t) = -\log S(t|\bfX) = -\log S_0(t \exp(\bfX^\top \bfbeta)).
\end{equation*}

Differentiating with respect to \( t \), we obtain the hazard function:
\begin{equation*}
	\begin{aligned}
		\Lambda_{\bfX}'(t) & = - \frac{\dif}{\dif t} \log S_0(t \exp(\bfX^\top \bfbeta))                                              \\
		                   & = - \frac{S_0'(t \exp(\bfX^\top \bfbeta)) \cdot \exp(\bfX^\top \bfbeta)}{S_0(t \exp(\bfX^\top \bfbeta))} \\
		                   & = h_0(t \exp(\bfX^\top \bfbeta)) \cdot \exp(\bfX^\top \bfbeta),
	\end{aligned}
\end{equation*}
where \( h_0(u) = -\frac{d}{du} \log S_0(u) \) is the baseline hazard function. Thus, the hazard function under the AFT model can be expressed as

\begin{equation}
	\Lambda_{\bfX}'(t) = q(\Lambda_{\bfX}(t)) \cdot \exp(\bfX^\top \bfbeta), \tag{4}
\end{equation}
where \( q(\cdot) \geq 0 \) is a function determined by the distribution of the error term \( \varepsilon \), characterizing the stochastic component of the AFT model within this ordinary differential equation framework.
